Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-29 paper/paper.tex
pass on section 4.1

The first step in the \tname's workflow is Reverse Engineering. The
goal of this step is to extract options from the malicious traffic. A
rule that contains these options, by construction, captures the
malicious traffic. \tname\ uses two strategies to extract options:
1)~field (for any protocol) and 2)~payload (for http only). We
elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the data in the fields of a network
message. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from
\suri. Section~\ref{sec:active-recon} shows a concrete example where
the values for options are directly extracted from packet fields. We
obtained the mapping of network fields to options manually, from the
\suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload; not the fields. We found empirically that this occurs
frequently in http attacks, which are prevalent. For example,
\Fix{\Luc{53\%}} of the \numrulessuri{} rules from a public data
set~\cite{emerging-threats-open} we analyzed prescribe defenses
against http attacks. Figure~\ref{fig:rule-jboss} shows an example of
a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  Content options are very
common in http rules---\percRulesWithContent{} of the rules from the
data set mentioned above contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the histogram of
the number of these options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  Consequently, handling content options
for http is important. The content strategy focuses on extracting such
kinds of options from the payload.

We found, inspecting existing rules, that \CodeIn{content} options
typically refer to sequence of characters in the payload. To find the
values associated with these options, \tname{} splits the payload of
the malicious message in tokens using natural language delimiters,
such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. Considering the example from
Section~\ref{sec:content-example}, \tname{} produced a total of 23
tokens, considered content option candidates.

>>>>

The first step in the \tname's workflow is Reverse Engineering. The
goal of this step is to extract options from the malicious
traffic. Consequently, a rule that contains these options captures the
malicious traffic. \tname\ uses different strategies to extract
options based on the protocol detected in the input traffic: 1)~the
field strategy applies to any protocol and 2)~the payload strategy
applies only to http. We elaborate these strategies in the following.

For the field strategy, \tname{} parses the input messages looking for
options associated with the network message field
data. Section~\ref{sec:rules-and-packets} shows the correspondence
between packet fields and some of the rule options from \suri and
Section~\ref{sec:active-recon} shows a concrete example where the
values for options are directly extracted from the fields of
messages. We obtained the option-field mapping manunually, from the
\suri\ documentation.

Unfortunately, the relevant data for the option may be at the message
payload; not the fields. We found empirically that this occurs
frequently in http attacks, which are prevalent. For example,
\percHttp\ of the \numrulessuri{} rules from a public data
set~\cite{emerging-threats-open} we analyzed prescribe defenses
against http attacks and the majority of these rules use content
options. Figure~\ref{fig:rule-jboss} shows an example of a rule that
captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
payload.  Content options are very common in http
rules---\percRulesWithContent{} of the rules from the data set
mentioned above contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the distribution
of number of content options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  To sum up, handling content options for
http is very important.

The content strategy focuses on extracting content options from the
message payload. Inspecting existing rules, we found that these
options refer, in the payload, to sequences of characters separated by
typical natural language delimiters such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. To find the values
associated with these options, \tname{} splits the payload of the
malicious message in tokens using these delimiters.  Considering the
example from Section~\ref{sec:content-example}, \tname{} produced a
total of 23 tokens, considered content option candidates.
